---
title: "CVTDM FINAL PROJECT - Lan Hoang and Monica Cruz"
output:
  html_document:
    df_print: paged
---

install.packages("lubridate")
install.packages("streamgraph")
```{r}
library(caret)
library(dummies)
library(leaps)
library(ggplot2)
library(forecast)
library(gplots)
library(FNN)
library(rpart)
library(rpart.plot)
library(dplyr)
library (plotly)
library(lubridate)
library(data.table)
library(viridis)
library(tidyverse)
library(DT)
```
 Reading datasets

```{r}
rm(list = ls()) # clean environment
cat("\014") # clean console
setwd("C:\\Users\\lanhh\\OneDrive\\Documents\\UNIGE Statistics\\Masters of Biz Analytics\\Creating Value through Data Mining\\Fall 2021\\5. Final Project")
```


```{r}
sales_train.df <- read.csv("sales_train.csv", header = T, sep = ",", na.strings=c("")) #importing data

sales_test.df <- read.csv("test.csv", header = T, sep = ",", na.strings=c(""))
```

```{r}
items.df <- read.csv("items.csv", header = T, sep = ",", na.strings=c(""))

item_categories.df <- read.csv("item_categories.csv", header = T, sep = ",", na.strings=c(""))

shops.df <- read.csv("shops.csv", header = T, sep = ",", na.strings=c(""))

```

Since the data set might have been designed for Python with the index starting from 0, we adjust the index to start from 1 in this test set to work with R.


```{r}
new_sales_train <- merge(sales_train.df, items.df, by = "item_id", all.x = TRUE)
#merge_item_shop <- merge(merge_item, shops.df, by = "shop_id", all.x = TRUE)

#merge_item_shop_itemcat <- merge(merge_item_shop, item_categories.df, by = "item_category_id", all.x = TRUE)
```

```{r}
#drop the columns with item_id, shop_id and item_category_id

#drops <- c("item_id","shop_id", "item_category_id")
#sales_train_merged <- merge_item_shop_itemcat[ , !(names(merge_item_shop_itemcat) %in% drops)]
```


```{r}
summary(new_sales_train)
```
No N/A or missing values in the sales_train set, which is a good thing


```{r}
#convert the date column from string to date format
new_sales_train$date <- as.Date(new_sales_train$date, "%d.%m.%Y")
```

```{r}
#attempt to reorder the data by date column, but the index column goes with it and is not reset according to the new order
#new <- sales_train_merged[order(sales_train_merged$date),]
```

```{r}
library(tidyverse)
```

```{r}
#reoder the dataframe by date and reset the index numbering according to the new order

#new_sales_train <- new_sales_train %>% arrange(date, item_id)
```

Finding the min and max date of the dataset

```{r}
min(new_sales_train$date)
max(new_sales_train$date)
summary(new_sales_train)
```
Now moving onto data cleaning. We notice that there are item prices that are less 
than 0, we need to remove that as it's not logical. 

```{r}
new_sales_train <- new_sales_train[new_sales_train$item_price > 0, ]
```

Data pre-processing

```{r}
#dropping some less useful features like item_name
new_sales_train$item_name <- NULL

```

```{r}
#sort data set by date
new_sales_train <- new_sales_train %>% arrange(date, item_id)

```


```{r}
#aggregate data by month (date_block_num)
data_group <- new_sales_train %>%                                 # Group data
  group_by(date_block_num, shop_id, item_category_id, item_id) %>%
  dplyr::summarize(sum_item_price = sum(item_price), mean_item_price = mean(item_price),
                   transaction = n(), sum_item_cnt = sum(item_cnt_day), mean_item_cnt = mean(item_cnt_day)) %>% 
  as.data.frame()
data_group   
```

```{r}
## Extract time based features.
data_group$year <- (data_group$date_block_num%/%12) + 2013
data_group$month <- (data_group$date_block_num%%12)

```

EXPLORATORY DATA ANALYSIS

Now need to visualize data, as well as check and remove any outliers (some treated 
"item_cnt" > 20 and < 0, "item_price" >= 400000 as outliers, so perhaps we could removed them as such,
but up to you)

```{r}
options(scipen = 999)
str(new_sales_train)
#adding the daily value of sales 
new_sales_train$amount<-c(new_sales_train$item_price*new_sales_train$item_cnt_day)

#converting daily data into monthly
set.seed(1)

new_sales_train$month<-floor_date(new_sales_train$date, "month")

monthly_mean<-new_sales_train%>%
                 group_by(month) %>%
                  summarize(mean = mean(amount))

monthly_dollars<-new_sales_train%>%
                  group_by(month) %>%
                  summarize(total.sales = sum(amount))


monthly_sales<- data.frame(c(monthly_dollars, monthly_mean))
monthly_sales
                            

#Plotting monthly sales in dollars
data<-monthly_dollars
plot_sales<-data %>%
  ggplot( aes(x=month, y=total.sales)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("monthly sales in dollars") 

plot_sales<-ggplotly(plot_sales)
plot_sales


#Plotting monthly mean of sales 
data<-monthly_mean
plot_meansales<-data %>%
  ggplot( aes(x=month, y=mean)) +
    geom_area(fill="#69b3a7", alpha=0.5) +
    geom_line(color="#69b3a3") +
    ylab("monthly mean of sales") 

plot_meansales<-ggplotly(plot_meansales)
plot_meansales

```

```{r}
options(scipen = 999)
#converting daily data into monthly for the item counts (how many items sold per month)
set.seed(1)
new_sales_train$month<-floor_date(new_sales_train$date, "month")

monthly_items<-new_sales_train%>%
                  group_by(month) %>%
                  summarize(total.items = sum(item_cnt_day))

monthly_items_mean<-new_sales_train%>%
                 group_by(month) %>%
                  summarize(mean = mean(item_cnt_day))


monthly_items.sales<- data.frame(c(monthly_items, monthly_items_mean))
monthly_items.sales

#Plotting monthly items sold 
data<-monthly_items
plot_monthly.items<-data %>%
  ggplot( aes(x=month, y=total.items)) +
    geom_area(fill="#69b3a7", alpha=0.5) +
    geom_line(color="#69b3a3") +
    ylab("total monthly items sold") 

plot_monthly.items<-ggplotly(plot_monthly.items)
plot_monthly.items

#Plotting monthly mean of items sold 
data<-monthly_items_mean
plot_monthly.items.mean<-data %>%
  ggplot( aes(x=month, y=mean)) +
    geom_area(fill="#69b3a7", alpha=0.5) +
    geom_line(color="#69b3a3") +
    ylab("mean monthly items sold") 

plot_monthly.items.mean<-ggplotly(plot_monthly.items.mean)
plot_monthly.items.mean

```

```{r}

#Monthly sales per shop
df.monthsSHOPsales = aggregate(list(Sales=new_sales_train$amount),by=list(shopID=(new_sales_train$shop_id), Month=cut(as.POSIXct(new_sales_train$date),"month")),sum)

df.monthsSHOPsales.fin<- df.monthsSHOPsales %>% mutate(id = row_number())
df.monthsSHOPsales.fin

#Plotting monthly sales per shop 
par=(cex=.1)
data <- df.monthsSHOPsales
data %>%
  ggplot( aes(x=Month, y=Sales, group=shopID, color=shopID)) +
    geom_line()


#splitting data frame per month 2013 
data_frame_january13<- df.monthsSHOPsales.fin[1:45,]
data_frame_january13

#Barplot january 2013 sales per shop 
par=(cex=.5)
 ggplot(data=data_frame_january13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+ 
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

 #splitting data frame feb 2013 
data_frame_feb13<- df.monthsSHOPsales.fin[46:91,]
data_frame_feb13

#Barplot feb 2013 sales per shop
par=(cex=.5)
 ggplot(data=data_frame_feb13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame march 2013 
data_frame_march13<- df.monthsSHOPsales.fin[92:137,]
data_frame_march13

par=(cex=.5)
 ggplot(data=data_frame_march13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

 #splitting data frame april 2013 
data_frame_april13<- df.monthsSHOPsales.fin[138:183,]
data_frame_april13

par=(cex=.5)
 ggplot(data=data_frame_april13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))


 #splitting data frame may 2013
data_frame_may13<- df.monthsSHOPsales.fin[184:228,]
data_frame_may13

par=(cex=.5)
 ggplot(data=data_frame_may13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame june 2013
data_frame_june13<- df.monthsSHOPsales.fin[229:274,]
data_frame_june13

par=(cex=.5)
 ggplot(data=data_frame_june13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame july 2013
data_frame_july13<- df.monthsSHOPsales.fin[275:320,]
data_frame_july13

par=(cex=.5)
 ggplot(data=data_frame_july13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame august 2013
data_frame_aug13<- df.monthsSHOPsales.fin[321:365,]
data_frame_aug13

par=(cex=.5)
 ggplot(data=data_frame_aug13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame september 2013
data_frame_sep13<- df.monthsSHOPsales.fin[366:410,]
data_frame_sep13

par=(cex=.5)
 ggplot(data=data_frame_sep13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame october 2013
data_frame_oct13<- df.monthsSHOPsales.fin[411:456,]
data_frame_oct13

par=(cex=.5)
 ggplot(data=data_frame_oct13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame november 2013
data_frame_nov13<- df.monthsSHOPsales.fin[457:501,]
data_frame_nov13

par=(cex=.5)
 ggplot(data=data_frame_nov13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame december 2013
data_frame_dec13<- df.monthsSHOPsales.fin[502:547,]
data_frame_dec13

par=(cex=.5)
 ggplot(data=data_frame_dec13, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

```

```{r}
#splitting data frame per month 2014
data_frame_january14<- df.monthsSHOPsales.fin[548:593,]
data_frame_january14

#Barplot january 2014 sales per shop 
par=(cex=.5)
 ggplot(data=data_frame_january14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

 #splitting data frame feb  
data_frame_feb14<- df.monthsSHOPsales.fin[594:639,]
data_frame_feb14

#Barplot feb 2014 sales per shop
par=(cex=.5)
 ggplot(data=data_frame_feb14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame march 2014 
data_frame_march14<- df.monthsSHOPsales.fin[640:687,]
data_frame_march14

par=(cex=.5)
 ggplot(data=data_frame_march14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

 #splitting data frame april 2014 
data_frame_april14<- df.monthsSHOPsales.fin[688:736,]
data_frame_april14

par=(cex=.5)
 ggplot(data=data_frame_april14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))


 #splitting data frame may 2014
data_frame_may14<- df.monthsSHOPsales.fin[737:785,]
data_frame_may14

par=(cex=.5)
 ggplot(data=data_frame_may14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame june 2014
data_frame_june14<- df.monthsSHOPsales.fin[786:834,]
data_frame_june14

par=(cex=.5)
 ggplot(data=data_frame_june14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame july 2014
data_frame_july14<- df.monthsSHOPsales.fin[835:884,]
data_frame_july14

par=(cex=.5)
 ggplot(data=data_frame_july14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame august 2014
data_frame_aug14<- df.monthsSHOPsales.fin[885:935,]
data_frame_aug14

par=(cex=.5)
 ggplot(data=data_frame_aug14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame september 2014
data_frame_sep14<- df.monthsSHOPsales.fin[936:985,]
data_frame_sep14

par=(cex=.5)
 ggplot(data=data_frame_sep14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))

#splitting data frame october 2014
data_frame_oct14<- df.monthsSHOPsales.fin[986:1000,]
data_frame_oct14

par=(cex=.5)
 ggplot(data=data_frame_oct14, aes(x=shopID, y=Sales))+
  xlab("Stores")+ ylab("Number of sales")+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(face="bold",color="black",size=8, angle=45))+
  theme(axis.text.y = element_text(face="bold",color="black",size=8))
```
```{r}

#Plotting Month vs. Item_cnt (sum)
sum_month.plot<- ggplot(data_group, aes(x=data_group$month, y=data_group$sum_item_cnt, color ="aquamarine2"))+
  geom_line()
sum_month.plot

  
#Plotting Month vs. Item_cnt (mean)
mean_month.plot<- ggplot(data_group, aes(x=data_group$month, y=data_group$mean_item_cnt, color ="aquamarine2"))+
  geom_line()
mean_month.plot 
  

#Plotting Mean quantity of Items by shop 
mean_item.plot<- ggplot(data_group, aes(x=data_group$shop_id, y=data_group$mean_item_cnt))+
  geom_bar(stat = "identity", fill = "#FF6666")

mean_item.plot

#Plotting Sum quantity of Items by shop 
sum_item.plot<- ggplot(data_group, aes(x=data_group$shop_id, y=data_group$sum_item_cnt))+
  geom_bar(stat = "identity", fill = "#FF6666") 

sum_item.plot



#Plotting Item_category_ID vs. Item_cnt (Sum)

sum_itemcat.plot<- ggplot(data_group, aes(x=data_group$item_category_id, y=data_group$sum_item_cnt))+
  geom_bar(stat = "identity", fill = "royalblue") 

sum_itemcat.plot

#Plotting Item_category_ID vs. Item_cnt (Mean)
mean_itemcat.plot<- ggplot(data_group, aes(x=data_group$item_category_id, y=data_group$mean_item_cnt))+
  geom_bar(stat = "identity", fill = "royalblue") 
mean_itemcat.plot

```
```{r}
#Plotting items quantity per month 
data_months<-data_group
plot_monthly.mean<-data_months %>%
  ggplot( aes(x=month, y=mean_item_cnt)) +
    geom_area(fill="#69b3a7", alpha=0.5) +
    geom_line(color="#69b3a3") +
    ylab("Mean item cnt") 

plot_monthly.mean<-ggplotly(plot_monthly.mean)
plot_monthly.mean
```






